---
title: "A data-driven approach to evaluating landslide hazards"
author: "Elsa Culler"
output:
  html_document:
    toc: yes
    df_print: paged
    code_folding: hide
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(include = TRUE, 
                      eval = TRUE,
                      message = FALSE,
                      warning = FALSE,
                      results = 'show',
                      cache = FALSE,
                      cache.lazy = FALSE,
                      fig.path = "figures/",
                      fig.retina = 1,
                      dpi = 450)
```

```{r libraries, include=FALSE}
# Analysis libraries
library(broom)
library(cluster)
library(cowplot)
library(furrr)
library(grid)
library(gtable)
library(lubridate)
library(maps)
library(scales)
library(sf)
library(tigris)
library(RColorBrewer)
library(zoo)
library(tidyverse)
```



# Data

This study uses landslide, fire, and precipitation data to study the 
antecedent precipitation of post-wildfire landslides. The impact of wildfire 
on rainfall-triggered landslide susceptibility in a variety of global regions 
is evaluated through a proxy of normalized precipitation.

```{r data_management, include = F}
clear.data <- function (variable.name, delete.rds = F) {
  if (exists(variable.name)) {
    rm(list = variable.name, envir = .GlobalEnv)
  }
  rds.path <- str_c('rds/', variable.name, '.Rds')
  if (file.exists(rds.path) && delete.rds) {
    #Delete file if it exists
    file.remove(rds.path)
  }
}

get.data <- function (variable.name, load.function, ..., 
                      reload.from.rds = F, overwrite = F) {
  rds.path <- str_c('rds/', variable.name, '.Rds')
  # Remove existing data if desired
  if (overwrite) {
    clear.data(variable.name, delete.rds = T)
  } 
  # Remove only environment variable to reload Rds file
  if (reload.from.rds) {
    clear.data(variable.name, delete.rds = F)
  }
  
  # If the variable doesn't exist (default)...
  if (!exists(variable.name)) { 
    # If the Rds file already exists...
    if (file.exists(rds.path)) { 
      # Read in the existing Rds file and assign the results to the variable
      assign(variable.name, readRDS(rds.path), envir = .GlobalEnv)
    # Otherwise it the Rds file doesn't exist
    } else {
      # Calculate the value from scratch and assign it to the variable...
      data <- load.function(...)
      assign(variable.name, data, envir = .GlobalEnv)
      # and if the user wants to overwrite the rds (default)...
      #store it to an Rds file
      saveRDS(data, rds.path)
    }
  }
}
```

## Data Filepaths

Use this section to modify file paths if running the notebook on a different
computer.
```{r data_paths, class.source = 'fold.show'}
# Landslide data from the NASA Global Landslide Catalog
slide.path <- '../regions_data/glc/GLC20180821.csv'

# Fire data from the MODIS Burned Area dataset
modis.paths <- list.files(path='../regions_data/processed', 
                          pattern='slide_modisburndate_[ns][ew].csv',
                          full.names = T)

# Precipitation data from CHIRPS
precip.chirps.ww.path <-  file.path("../regions_data/processed", 
                                    "glc_precip_global.csv")
precip.chirps.ww.monthly.path <- file.path("../regions_data/", "processed",
                                   "glc_precip_global_monthly.csv")

# Additional preciptiation data from Daymet for a limited region
precip.daymet.ca.path <- file.path("../regions_data/processed", 
                                   "glc_precip_daymet.csv")
```

## Data inclusion

Include landslides with the following parameters:
* Rainfall triggered
```{r data_parameters_rainfall, class.source = 'fold-show'}
rain.triggers <- c('rain', 'downpour', 'flooding', 'continuous_rain')
```

* Occurred within the time frame of the fire and precipitation datasets
```{r data_parameters_dates, class.source = 'fold-show'}
start.date <- lubridate::ymd('2000-01-01')
end.date <- lubridate::ymd('2019-01-01')
```

* Occurred within the spatial extend of the fire and precipitation datasets
```{r data_parameters_spatial, class.source = 'fold-show'}
min.latitude <- -50
max.latitude <- 50
```

* Location accuracy on par with precipitation resolution or more accurate
```{r data_parameters_accuracy, class.source = 'fold-show'}
min.location.accuracy <- '10km'
```

* MODIS Burned Area data should include an additional 3 years
```{r data_parameters_modis, class.source = 'fold-show'}
modis.start.date <- lubridate::ymd('20040101')
modis.end.date <- lubridate::ymd('20190101')
```

* Wilcox tests between burned and unburned locations have the following parameters:
```{r wilcox_parameters, class.source = 'fold-show'}
wilcox.alternative <- 'less' # n - burned, m - unburned
wilcox.alpha <- 0.05
```

## Study definitions

- Study location parameters are applied using a set of functions to identify if data
are included in the study.
```{r location_inclusion_functions}
# Identify if a location is in the global limits
is.in.global.study <- function (x, y) { between(y, min.latitude, max.latitude) }

# Identify if a location is in the CA/NV pilot study
get.data('state.boundaries', tigris::states)
boundaries <- subset(state.boundaries, STUSPS %in% c('CA', 'NV'))
boundaries.df <- fortify(boundaries)
is.in.ca.study <- function (x, y) {
  bbox <- sf::st_bbox(subset(state.boundaries, STUSPS == 'CA'))
  between(x, bbox$xmin, bbox$xmax) &
    between(y, bbox$ymin, bbox$ymax)
}
```

- Seasons are split in the Northern Hemisphere by:

    Winter
    : December-January-February
    
    Spring
    : March-April-May
    
    Summer
    : June-July-August
    
    Fall
    : September-October-November

- In the southern hemisphere, the split is shifted instead:

    Winter
    : June-July-August
    
    Spring
    : September-October-November
    
    Summer
    : December-January-February
    
    Fall
    : March-April-May

- A set of functions converts months and days-of-the-year to seasons.
```{r season_functions}
season.breaks <- c('spring'=59, 'summer'=151, 'fall'=243, 'winter'=334)
yday.to.season <- function(yday, latitude) {
    if (latitude > 0) {
      case_when(
        yday > season.breaks['winter'] ~ 'Winter',
        yday > season.breaks['fall'] ~ 'Fall',
        yday > season.breaks['summer'] ~ 'Summer',
        yday > season.breaks['spring'] ~ 'Spring',
        TRUE ~ 'Winter')
    } else {
      case_when(
        yday > season.breaks['winter'] ~ 'Summer',
        yday > season.breaks['fall'] ~ 'Spring',
        yday > season.breaks['summer'] ~ 'Winter',
        yday > season.breaks['spring'] ~ 'Fall',
        TRUE ~ 'Summer')
    }
}

month.to.season <- function (month, north) {
  if (north) {
    case_when(
      month >= 12 ~ 'Winter',
      month >= 9 ~ 'Fall',
      month >= 6 ~ 'Summer',
      month >= 3 ~ 'Spring',
      TRUE ~ 'Winter')
  } else {
    case_when(
      month >= 12 ~ 'Summer',
      month >= 9 ~ 'Spring',
      month >= 6 ~ 'Winter',
      month >= 3 ~ 'Fall',
      TRUE ~ 'Summer')
  }
}

```
- Consistent plot colors and themes
```{r plot_colors}
# Region colors
region.plot.order <- c('All landslides',
                       'California',
                       'Intermountain West',
                       'Pacific Northwest',
                       'Central America',
                       'Himalayas',
                       'Southeast Asia')
scale_region_values <- c('All landslides' = 'grey50',
                         setNames(RColorBrewer::brewer.pal(6, 'Dark2'), 
                                  region.plot.order[2:7]))
scale_region_labels <- c('All landslides' = 'None',
                         setNames(region.plot.order[2:7], 
                                  region.plot.order[2:7]))
scale_color_region <- scale_color_manual('Region', 
                                         values = scale_region_values,
                                         labels = scale_region_labels)
scale_fill_region <- scale_fill_manual('Region', 
                                       values = scale_region_values,
                                       labels = scale_region_labels)

scale_color_precip <- scale_color_distiller('Precipitation percentile', 
                                            palette='BrBG', direction = 1,
                                            guide = guide_colorbar(barwidth = unit(0.5, 'npc')))
scale_fill_precip <- scale_fill_distiller('Precipitation percentile', 
                                          palette='BrBG', direction = 1,
                                          guide = guide_colorbar(barwidth = unit(0.5, 'npc')))

# Fire colors
burned.labels <- c('burned'='Burned', 'unburned'='Unburned')
burned.colors <- c('burned' = '#d95f02', unburned = '#7570b3')
scale_color_burned <- scale_color_manual('', 
                                         values = burned.colors,
                                         labels = burned.labels)
scale_fill_burned <- scale_fill_manual('', 
                                       values = burned.colors,
                                       labels = burned.labels) 
# Season
scale_x_season <- scale_x_continuous(
    breaks = season.breaks,
    labels = c('Spring', 'Summer', 'Fall', 'Winter'))

col.ocean <- 'lightskyblue1'
```

```{r plot_themes}
theme_map <- list(
  theme_bw(),
  coord_quickmap(),
  labs(x='', y='')
)

theme_global_map <- list(theme_map,
                         borders(fill = 'grey90', size = 0.1),
                         theme(panel.background = element_rect(fill=col.ocean), 
                               panel.grid = element_blank(),
                               axis.text = element_blank(),
                               axis.ticks = element_blank(),
                               plot.margin = unit(c(0,0,0,0), "null")),
                         coord_quickmap(ylim = c(-60, 75), xlim = c(-180, 180))
)
theme_panel <- theme(panel.background = element_rect(fill=col.ocean), 
                     panel.grid = element_blank(),
                     axis.text = element_blank(),
                     axis.ticks = element_blank(),
                     plot.margin = unit(c(1,0,-1,0), "lines"),
                     legend.position = 'null')
theme_panel_legend <- theme(
  legend.box.margin = margin(-1, 0, 1, 0),
  legend.box.just = c(0, 0),
  #legend.background = element_rect(color='black'), #debug
  legend.position = 'bottom'
)
capitalize <- function(string) {
  substr(string, 1, 1) <- toupper(substr(string, 1, 1))
  string
}
```

## Landslide data from the NASA Global Landslide Catalog

```{r load_slides}
load.slides <- function (slide.path, 
                         start.date, end.date, 
                         is.in.global.study, 
                         rain.triggers, 
                         min.location.accuracy) {
  # Define factor levels
  ls.sizes <- c("small", "medium", "large", "very_large", "catastrophic")
  ls.accuracy <- c("exact", "1km", "5km", "10km", 
                   "25km", "50km", "100km", "250km")
  
  # Read data file and remove duplicates
  slide.raw <- readr::read_csv(slide.path,
    guess_max = 100000,
    na = c("", "unknown", "--"),
    col_types = cols(
      landslide_size = readr::col_factor(ls.sizes, ordered=TRUE),
      location_accuracy = readr::col_factor(ls.accuracy, ordered=TRUE),
      landslide_trigger = readr::col_factor(NULL),
      landslide_category = readr::col_factor(NULL),
      landslide_setting = readr::col_factor(NULL),
      country_code = readr::col_factor(NULL)
    )
  ) %>% 
    dplyr::mutate(event_date = as_date(ymd_hm(event_date)),
                  submitted_date = as_date(ymd_hm(submitted_date)),
                  last_edited_date = as_date(ymd_hm(last_edited_date)))
  slide.unique <- slide.raw  %>%
    dplyr::group_by(event_date, latitude, longitude) %>%
    dplyr::mutate(i = row_number()) %>%
    dplyr::filter(i == 1) %>%
    dplyr::select(-i) %>%
    dplyr::ungroup()
  
  # Filter landslides to meet study parameters
  slide.unique %>%
    dplyr::filter(event_date >= start.date,
                  event_date <= end.date,
                  is.in.global.study(longitude, latitude),
                  landslide_trigger %in% rain.triggers,
                  location_accuracy <= min.location.accuracy) %>%
    droplevels() %>%
    dplyr::mutate(event_yday = yday(event_date),
                  debris_flow = landslide_category %in% c('mudslide', 'debris flow'),
                  season = factor(map2_chr(event_yday, latitude, 
                                           ~yday.to.season(.x, .y)),
                                  levels = c('Spring', 'Summer', 'Fall', 'Winter'),
                                  ordered = T))
}

get.data('slide.all.precip', overwrite = F,
         load.slides,
         slide.path, start.date, end.date, 
         is.in.global.study, rain.triggers,
         min.location.accuracy)

slide.all.precip %>%
  dplyr::select(landslide_trigger, location_accuracy, latitude, event_date) %>%
  summary()
```
## Precipitation data from CHIRPS and Daymet

### Global CHIRPS data
```{r load_chirps_all}
load.precip.all <- function (precip.chirps.ww.path) {
    precip.chirps.ww.raw <- read_csv(precip.chirps.ww.path,
                                     col_types = cols(X1 = col_datetime(format = "%Y-%m-%d")))
    precip.chirps.ww.raw %>%
      dplyr::select(date=X1, OBJECTID, pctl=precip_pctl_7) %>%
      dplyr::filter(OBJECTID %in% slide$OBJECTID) %>%
      dplyr::mutate(window=7) %>%
      dplyr::filter(!is.na(pctl)) # Remove NA values from beginning of data range
}

get.data('precip.chirps.ww.all', precip.chirps.ww.path)
precip.chirps.ww.all
```

### Detailed global CHIRPS data close to the landslide
```{r load_chirps_event}
date.filter.precip <- function (raw, slide=slide.all.precip, pre=90, post=30) {
  raw %>%
    dplyr::right_join(slide %>% dplyr::select(OBJECTID, event_date), 
                      by='OBJECTID') %>% 
    dplyr::select(OBJECTID, dplyr::starts_with('precip_'), event_date, date) %>%
    dplyr::mutate(date = lubridate::as_date(date),
           days_before = as.integer(event_date - date)) %>%
    dplyr::filter(days_before < pre, days_before > -post)
}

tidy.precip <- function (precip) {
  precip %>% 
    dplyr::select(OBJECTID, starts_with('precip_'), event_date, date) %>%
    gather(key='variable', value='amount', starts_with('precip_')) %>%
    separate(variable, into=c(NA, 'variable', 'window')) %>% 
    dplyr::filter(variable=='pctl') %>%
    dplyr::select(-variable, pctl=amount) %>%
    dplyr::mutate(days_before = as.integer(event_date - date),
           yday = yday(event_date - days(days_before)),
           window = as.numeric(window)) %>%
    dplyr::select(-event_date)
}

load.event.precip <- function (precip.chirps.ww.path) {
    precip.chirps.ww.raw <- read_csv(precip.chirps.ww.path,
                                     col_types  = readr::cols(X1 = col_datetime(format = "%Y-%m-%d")))
    precip.chirps.ww.raw.filt <- precip.chirps.ww.raw %>%
      dplyr::rename(date = X1) %>%
      dplyr::filter(year(date) > 2003)
    precip.chirps.ww.raw.filt2 <- precip.chirps.ww.raw.filt %>%
      date.filter.precip 
    precip.chirps.ww.raw.filt2 %>% 
      tidy.precip
}

get.data('precip.chirps.ww.event', precip.chirps.ww.path)
precip.chirps.ww.event
```

### Import Monthly CHIRPS data
```{r load_chirps_monthly}
load.precip.monthly <- function (monthly.path, slide) {
  read_csv(precip.chirps.ww.monthly.path,
           col_types = cols(X1 = col_date(format = '%Y-%m-%d'))) %>%
    dplyr::transmute(OBJECTID, 
                     year = year(X1), month = month(X1), 
                     precip.mm=precip) %>%
    dplyr::inner_join(slide %>% select(OBJECTID, latitude), by='OBJECTID') %>%
    dplyr::mutate(north = latitude > 0) %>%
    dplyr::group_by(north) %>%
    dplyr::mutate(season = factor(month.to.season(month, north),
                                  levels = c('Spring', 'Summer', 'Fall', 'Winter'),
                                  ordered = T)) %>%
    ungroup()
}
 
get.data('precip.chirps.ww.monthly', overwrite = F,
         load.precip.monthly,
         precip.chirps.ww.monthly.path,
         slide.all.precip)

precip.chirps.ww.seasonal.median <- precip.chirps.ww.monthly %>%
  group_by(OBJECTID, year, season) %>%
  summarize(precip.mm = sum(precip.mm)) %>%
  ungroup() %>%
  summarize(median = median(precip.mm)) %>%
  pull(median)

precip.chirps.ww.monthly
```


### Filter out landslides with zero precipitation
```{r tally_excluded}
precip.chirps.ww.event %>%
  # Filter out sites with no precipitation data at all
  filter(days_before==-1, window==7, !is.na(pctl)) %>%
  # Filter out sites already excluded by landslide criteria
  inner_join(slide.all.precip, by = 'OBJECTID') %>%
  # Count numbers of zero and non-zero precipitation
  group_by(precip = ifelse(pctl > 0, 'non.zero.precip', 'zero.precip')) %>%
  tally() %>%
  spread(key = precip, value = n) %>%
  mutate(total = non.zero.precip + zero.precip,
         percent.zero.precip = zero.precip / total)
```

```{r filter_slide_no_precip}
filter.slide.precip <- function (slide.all.precip, precip.event) {
  slide.all.precip %>%
    dplyr::inner_join(precip.event %>% 
                        filter(days_before==-1, window==7, !is.na(pctl)) %>%
                        dplyr::filter(days_before==-1, window==7, pctl > 0)) %>%
    dplyr::select(OBJECTID, latitude, longitude, event_date, event_yday, season,
                  location_accuracy, 
                  landslide_category, landslide_trigger, landslide_size,
                  landslide_setting, debris_flow)
}
get.data('slide',  overwrite = F,
         filter.slide.precip,
         slide.all.precip,
         precip.chirps.ww.event)
slide
```

### Filter California Landslides
```{r slide_california}
filter.slide.ca <- function (slide.all.precip) {
  slide.all.precip %>%
    filter(is.in.ca.study(longitude, latitude)) %>%
    mutate(region = as.factor(case_when(
             latitude < 35.0 ~ 'South',
             longitude < -121.25 ~ 'North',
             TRUE ~ 'East')))
}

get.data('slide.ca',  overwrite = F,
         filter.slide.ca, 
         slide.all.precip)

slide %>% summary
```

### Filter California CHIRPS data
```{r chirps_precip_ca}
filter.precip.ca <- function (precip.chirps.ww.event, slide.ca) {
  precip.chirps.ca <- precip.chirps.ww.event%>%
    dplyr::filter(OBJECTID %in% slide.ca$OBJECTID)
}

get.data('precip.chirps.ca.event', filter.precip.ca,  overwrite = F,
         precip.chirps.ww.event, slide.ca)
get.data('precip.chirps.ca.all', filter.precip.ca, overwrite = F,
         precip.chirps.ww.all, slide.ca)

precip.chirps.ca.event %>%
  # Filter out sites with no precipitation data at all
  filter(days_before==-1, window==7, !is.na(pctl)) %>%
  # Filter out sites already excluded by landslide criteria
  inner_join(slide.all.precip, by = 'OBJECTID') %>%
  # Count numbers of zero and non-zero precipitation
  group_by(precip = ifelse(pctl > 0, 'non.zero.precip', 'zero.precip')) %>%
  tally() %>%
  spread(key = precip, value = n) %>%
  mutate(total = non.zero.precip + zero.precip,
         percent.zero.precip = zero.precip / total * 100)
```

### Import Daymet data for California
```{r dayment_precip.ca}
load.precip.daymet.ca <- function (precip.daymet.ca.path) {
  read_csv(precip.daymet.ca.path, 
           col_types = cols(X1 = col_datetime(format = "%Y-%m-%d %H:%M:%S"))) %>% 
  rename(date=X1) %>%
  date.filter.precip %>%
  tidy.precip
}

get.data('precip.daymet.ca.all', overwrite = F,
         load.precip.daymet.ca, 
         precip.daymet.ca.path)

precip.daymet.ca.all
```

## Fire data computed from MODIS Burned Area

```{r modis_find_fires}
load.modis <- function (modis.paths, burn.date.min, burn.date.max) {
  modis.date.raw <- dplyr::bind_rows(map(modis.paths, read_csv))
  burn.record.length <- time_length(burn.date.max - burn.date.min, 'years')
  modis.date.raw %>% 
    dplyr::right_join(slide %>% dplyr::select(OBJECTID, event_date, latitude), 
                      by='OBJECTID') %>%
    # Convert MODIS Burned Area value to date
    dplyr::mutate(burn_date = make_date(year=floor(month / 1000)) + burn_date) %>%
    dplyr::select(-month) %>%
    # Separate out indivicual fires
    dplyr::group_by(OBJECTID) %>%
    dplyr::mutate(gap = as.integer(burn_date - lag(burn_date, default=min(burn_date))),
                  new.fire = gap > 30,
                  fire.id = cumsum(new.fire)) %>%
    drop_na(fire.id) %>%
    # Calculate burn fraction, fire ending date, and season for each fire
    dplyr::group_by(OBJECTID, event_date, fire.id) %>%
    summarize(fraction = sum(fraction), 
              burn_date = max(burn_date),
              burn_season = yday.to.season(yday(burn_date), latitude)) %>%
    # Calculate time relative to landslide
    dplyr::mutate(burn_days_before = event_date - burn_date,
                  burn_years_before = time_length(burn_days_before, 'years')) %>%
    # Compute fire information
    dplyr::group_by(OBJECTID) %>%
    dplyr::mutate(pre.slide = burn_date==max(burn_date[burn_days_before>=0]),
                  n = n(),
                  period = burn.record.length / n,
                  start_days_before = min(event_date) - burn.date.min,
                  start_years_before = time_length(start_days_before, 'years'),
                  end_days_before = max(event_date) - burn.date.max,
                  end_years_before = time_length(end_days_before, 'years')) %>%
    dplyr::select(-event_date) %>%
    dplyr::ungroup()
}

get.data('modis.date', overwrite = F,
         load.modis,
         modis.paths, modis.start.date, modis.end.date)
modis.date
```

```{r classify_burned}
classify.burned <- function (modis.date, slide) {
  modis.date  %>%
    dplyr::filter(burn_years_before < 3, pre.slide) %>%
    dplyr::right_join(slide, by='OBJECTID') %>%
    dplyr::group_by(OBJECTID) %>%
    summarize(burn.cumulative = sum(fraction)) %>%
    dplyr::mutate(burn.cumulative = ifelse(is.na(burn.cumulative), 
                                           0, 
                                           burn.cumulative),
                  burned = ifelse(burn.cumulative > 0, 'burned', 'unburned')) %>%
    dplyr::select(OBJECTID, burned, burn.cumulative) %>% 
    dplyr::ungroup()
}

get.data('modis', overwrite = F,
         classify.burned, 
         modis.date,
         slide)

modis %>%
  group_by(burned) %>%
  tally() %>%
  mutate(percent = n / nrow(slide) * 100)
```

## Define regions using AGNES clustering

```{r cluster_regions, fig.width=16, fig.height=8, fig.show='hide'}
cluster.global <- function (slide, agclusts, k=30) {
  # Select clusters with more than 100 landslides, and label
  slide %>%
    dplyr::select(OBJECTID, latitude, longitude) %>%
    dplyr::mutate(region = agclusts %>% cutree(k)) %>%
    dplyr::group_by(region) %>%
    mutate(region = as_factor(region)) %>%
    dplyr::filter(n() > 100)  %>%
    dplyr::ungroup() %>%
    dplyr::mutate(region = fct_recode(as_factor(region),
                              'Central Europe' = '1', 
                              'Malaysia' = '2', 
                              'Western US/Canada' = '3',
                              'East Brazil' = '4',
                              'Southeast Asia' = '5',
                              'South India' = '6',
                              'Eastern US' = '8',
                              'Himalayas' = '9',
                              'Southeast Asia' = '10',
                              'Caribbean/Venezuela' = '11', 
                              'Central America' = '14'))
}

# Cluster landslides based on location using AGNES algorithm
get.data('agclusts', 
         cluster::agnes, 
         slide %>%
           dplyr::select(latitude, longitude))


get.data('clusters.global',  overwrite = F,
         cluster.global, 
         slide.all.precip, agclusts)

clusters.global %>%
    ggplot(aes(longitude, latitude)) +
    theme_global_map +
    geom_point(aes(color = region), size=1) +
    scale_color_brewer('Region', palette='Paired') +
    guides(color=guide_legend(override.aes = list(size=5)))
```

```{r split_us_regions, fig.show='hide'}
split.regions.us <- function (agclusts, slide, k=6) {
  # Crop cluster tree, select clusters with more than 100 landslides, and label
  slide %>%
    dplyr::select(OBJECTID, latitude, longitude) %>%
    dplyr::mutate(region = agclusts %>% cutree(k)) %>%
    mutate(region = as_factor(region)) %>%
    dplyr::mutate(region = fct_recode(as_factor(region),
                                      'Intermountain West' = '3',
                                      'Intermountain West' = '4',
                                      'Intermountain West' = '5',
                                      'Pacific Northwest' = '1',
                                      'California' = '2')) %>%
    dplyr::group_by(region) %>%
    dplyr::filter(n() > 100)  %>%
    dplyr::ungroup() %>%
    droplevels()
}

get.data('agclusts.us',  overwrite = F,
         function (coord) cluster::agnes(coord, stand=T), 
         clusters.global %>%
           dplyr::filter(region=='Western US/Canada') %>%
           dplyr::select(latitude, longitude))

get.data('clusters.us',  overwrite = F,
         split.regions.us,
         agclusts.us, 
         clusters.global %>%
           dplyr::filter(region=='Western US/Canada'))

clusters.us %>%
  ggplot(aes(longitude, latitude)) +
  geom_point(aes(color = region), size=1) +
  scale_color_brewer(palette='Set2') +
  guides(color=guide_legend(override.aes = list(size=5))) +
  theme_map
```

```{r merge_regions}
merge.regions <- function (clusters.global, clusters.us) {
  # Merge US and global regions
  clusters.global %>%
    left_join(clusters.us %>% dplyr::select(OBJECTID, region), 
              by='OBJECTID') %>%
    dplyr::mutate(region = as_factor(ifelse(!is.na(region.y), 
                                            as.character(region.y), 
                                            as.character(region.x)))) %>%
    dplyr::select(-starts_with('region.')) %>%
    filter(region != 'Western US/Canada') %>%
    droplevels()
}

get.data('regions.all', overwrite = F,
         merge.regions, 
         clusters.global, clusters.us)
```

```{r get_regions_info}
get.regions.info <- function (regions.all, slide) {
  regions.all %>%
    bind_rows(slide %>% dplyr::mutate(region = 'All landslides')) %>%
    dplyr::inner_join(modis, by = 'OBJECTID') %>%
    group_by(region) %>%
    dplyr::group_by(region, burned) %>%
    dplyr::summarize(n = n()) %>%
    tidyr::spread(key = burned, value = n) %>%
    dplyr::mutate(n = burned + unburned,
                  percent.burned = burned / n * 100) %>%
    ungroup
}

get.data('regions.info', overwrite = F,
         get.regions.info,
         regions.all, slide)

regions.info
```



```{r filter_and_merge_regions}
filter.and.merge.regions <- function (regions.all, regions.info, slide) {
  # Combine nearby regions
  regions.all  %>%
    filter(OBJECTID %in% slide$OBJECTID) %>%
    bind_rows(slide %>% dplyr::mutate(region = 'All landslides')) %>% 
    dplyr::left_join(regions.info) %>% 
    mutate(region = fct_recode(region,
                               'Central America' = 'Caribbean/Venezuela')) %>%
    dplyr::filter(region != 'East Brazil', 
                  percent.burned > 4) %>%
    droplevels() %>%
    dplyr::mutate(region = factor(region, region.plot.order, ordered = T)) %>%
    dplyr::select(OBJECTID, region)
}

get.data('slide.regions',  overwrite = F,
         filter.and.merge.regions, 
         regions.all, 
         regions.info, 
         slide)

regions.tally <- slide.regions %>%
  group_by(region) %>%
  tally

regions.tally
```


# Exploration
## Map

A panelled maps showing landslide regions (a and b), 
whether or not each landslide took place on a burned site (c and d), 
and the precipitation percentile of the landslide-triggering storm (e and f).
Inset plots show, for each region,
the distribution of burned fractions of the landslide sites (a and b),
location accuracy of landslides split by fire history (c and d), and the
seasonal total precipitation climatology (e and f).
### Regions map
```{r map_regions, fig.height=3, fig.width=6.5, fig.show='hide'}
gg.region <- slide.regions %>%
  left_join(slide, by='OBJECTID') %>%
  mutate(region = fct_relevel(region, 'All landslides', after = Inf)) %>%
  arrange(desc(region)) %>%
  ggplot(aes(longitude, latitude)) +
  theme_global_map +
  geom_point(aes(color = region), size=0.02) + 
  labs(color = '') +
  scale_color_region +
  guides(color = guide_legend(override.aes = list(size = 2))) +
  theme_panel_legend

gg.region
```

### Burned sites map
```{r map_fire, fig.height=3, fig.width=6.5, fig.show='hide'}
gg.burned <- slide %>%
  dplyr::left_join(modis) %>%
  dplyr::arrange(desc(burned)) %>%
  ggplot() +
  theme_global_map +
  #geom_point(data = slide.regions %>%
  #             dplyr::filter(region !='All landslides') %>%
  #             dplyr::left_join(slide, by='OBJECTID'),
  #           aes(x = longitude, y = latitude, fill = region),
  #           size = 2, alpha = .01, shape = 21, stroke = NA) +
  geom_point(aes(x=longitude, y=latitude, color=burned),
             size=.02) +
  scale_color_burned +
  #scale_fill_region +
  guides(color = guide_legend(override.aes = list(size = 2),
                              title = NULL),
         fill = guide_none()) +
  theme_panel_legend

gg.burned
```

### Precipitation map
```{r map_precipitation, fig.height=3.5, fig.width=6.5, fig.show='hide'}
gg.precip <- slide %>%
  left_join(precip.chirps.ww.event %>% 
              filter(days_before==-1, window==7), 
            by='OBJECTID') %>%
  filter(pctl > 0) %>%
  arrange(desc(pctl)) %>%
  ggplot() +
  theme_global_map +
  geom_point(aes(x=longitude, y=latitude, color=pctl),
             size=.02) +
  scale_color_precip +
  theme_panel_legend

gg.precip
```

### Burned fraction

```{r fig.width=7, fig.height=1.05, fig.show='hide'}
min.burn.cumulative <- modis %>%
  filter(burn.cumulative > 0) %>%
  pull(burn.cumulative) %>%
  min
max.burn.cumulative <- modis %>%
  pull(burn.cumulative) %>%
  max
plot.burned.fraction <- function (reg, slide.regions, modis) {
  slide.regions %>%
    filter(region==reg) %>%
    left_join(modis, by = 'OBJECTID') %>%
    filter(burn.cumulative > 0) %>%
    ggplot() +
    geom_density(aes(x = burn.cumulative), fill = 'black') +
    scale_x_log10() +
    labs(x = 'Burned fraction', y = '') +
    scale_color_region +
    scale_fill_region +
    coord_cartesian(expand=F, xlim = c(min.burn.cumulative, max.burn.cumulative)) +
    theme_bw() +
    theme(legend.position = 'none',
          axis.text.x = element_text(angle = 90, vjust = 0.5, size = 4),
          axis.title = element_blank(),
          axis.text.y = element_blank(),
          axis.ticks.y = element_blank(),
          plot.background = element_blank(),
          panel.grid = element_blank(),
          plot.title = element_text(size = 5, margin=margin(0,0,0,0))) +
    ggtitle(str_c(reg, '\nn=', pull(filter(regions.tally, region==reg))))
}

gg.burned.fraction <- region.plot.order %>%
  set_names %>%
  map(plot.burned.fraction, slide.regions, modis)

cowplot::plot_grid(plotlist = gg.burned.fraction, nrow = 1)
```

### Location accuracy by region/fire history
```{r accuracy_by_region, fig.width=7, fig.height=1.05, fig.show='hide'}
plot.accuracy <- function (reg, slide.regions, slide, modis) {
  slide.regions %>%
    left_join(slide, by='OBJECTID') %>%
    left_join(modis, by='OBJECTID') %>%
    filter(region==reg) %>%
    mutate(location_accuracy = fct_recode(location_accuracy, '0km' = 'exact')) %>%
    drop_na(location_accuracy) %>%
    ggplot() +
    geom_bar(aes(x = location_accuracy, fill=burned,
                 group = interaction(region, burned),
                 y = ..prop..), 
             position=position_dodge2(preserve = 'single')) +
    scale_fill_burned +
    coord_cartesian(expand=F, ylim = c(0, 0.6)) +
    theme_bw() +
    labs(y = '', x = '', fill = 'Location\naccuracy') +
    theme(legend.position = 'none',
          axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1, 
                                     size = 4),
          axis.title = element_blank(),
          axis.text.y = element_blank(),
          axis.ticks.y = element_blank(),
          plot.background = element_blank(),
          panel.grid = element_blank(),
          plot.title = element_text(size = 5, margin=margin(0,0,0,0))) +
    ggtitle(str_c(reg, '\nn=', pull(filter(regions.tally, region==reg))))
}

gg.accuracy <- region.plot.order %>%
  set_names %>%
  map(plot.accuracy, slide.regions, slide, modis)

cowplot::plot_grid(plotlist = gg.accuracy, nrow = 1)
```

### Climatology by region
```{r climatology_by_region, fig.width=7, fig.height=1.05, fig.show='hide'}
plot.climatology <- function (reg, precip.monthly, slide.regions, med) {
  precip.seasonal <- precip.monthly %>%
    group_by(OBJECTID, year, season) %>%
    summarize(precip.mm = sum(precip.mm)) 
  
  ymin <- 1
  ymax <- max(precip.seasonal$precip.mm, na.rm=T)
  
  precip.seasonal %>%
    dplyr::right_join(slide.regions %>% filter(region==reg)) %>%
    drop_na(season) %>%
    ggplot()  +
    geom_violin(aes(y = precip.mm, x = season, group = season),
                fill='black', color=NA, adjust = 2)  +
    geom_hline(aes(yintercept = med), color = 'grey', size = .3)+
    scale_y_log10(breaks = c(1, 10, 100, 1000)) +
    scale_x_discrete(labels = c('Spring' = 'MAM', 'Summer' = 'JJA',
                                'Fall' = 'SON', 'Winter' = 'DJF'),
                     expand = ggplot2::expand_scale(add = c(1, 0))) +
    coord_cartesian(expand=F, ylim = c(ymin, ymax)) +
    theme_bw() +
    theme(legend.position = 'none',
          axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1, size = 4, 
                                     margin=margin(0,0,0,0)),
          axis.text.y = element_text(size = 4, hjust = 1, vjust = 0.5),
          axis.title = element_blank(),
          plot.background = element_blank(),
          panel.grid = element_blank(),
          plot.title = element_text(size = 5, margin=margin(0,0,0,0))) +
    ggtitle(str_c(reg, '\nn=', pull(filter(regions.tally, region==reg))))
}

gg.climatology <- region.plot.order %>%
  purrr::set_names() %>%
  purrr::map(plot.climatology, 
             precip.chirps.ww.monthly, 
             slide.regions, 
             precip.chirps.ww.seasonal.median)
 
cowplot::plot_grid(plotlist = gg.climatology, nrow = 1)
```

### Full map figure

```{r map_combined}
coord_america <- coord_equal(xlim = c(-140, -50), ylim = c(-8, 53), expand = F)
coord_asia <-  coord_equal(xlim = c(50, 140), ylim = c(-12, 49), expand = F) 

# Clip maps
gg.america.region <- gg.region + theme_panel + coord_america
gg.asia.region <- gg.region + theme_panel + coord_asia
gg.america.burned <- gg.burned + theme_panel + coord_america
gg.asia.burned <- gg.burned + theme_panel + coord_asia
gg.america.precip <- gg.precip + theme_panel + coord_america
gg.asia.precip <- gg.precip + theme_panel + coord_asia

# Pull legends
legend.regions <- cowplot::get_legend(gg.region)
legend.burned <- cowplot::get_legend(gg.burned)
legend.precip <- cowplot::get_legend(gg.precip)

# Region midpoints for arrows
region.centroids <- slide.regions %>%
  left_join(slide) %>%
  group_by(region) %>%
  summarize(y = mean(latitude), x = mean(longitude))
```

```{r insets_regions, fig.height=1.05, fig.width = 1.75, fig.show='hide'}

# Add insets
inset.w <- .2
inset.h <- .41
inset.arrow <- grid::arrow(length = unit(0.03, "npc"), type = 'open')
# Regions
gg.america.region.inset <- cowplot::ggdraw(gg.america.region) +
  # Arrows
  cowplot::draw_grob(grid::grid.lines(x = c(.1, .25), y = c(.8, .8), 
                                      arrow=inset.arrow)) +
  cowplot::draw_grob(grid::grid.lines(x = c(.1, .275), y = c(.4, .65), 
                                      arrow=inset.arrow)) +
  cowplot::draw_grob(grid::grid.lines(x = c(.4, .4), y = c(.375, .575), 
                                      arrow=inset.arrow)) +
  cowplot::draw_grob(grid::grid.lines(x = c(.825, .675), y = c(.675, .3), 
                                      arrow=inset.arrow)) +
  # Inset plots
  cowplot::draw_plot(gg.accuracy$`Pacific Northwest`, 
                     x = 0.01, y = 1, 
                     width = inset.w, height = inset.h, vjust = 1) +
  cowplot::draw_plot(gg.accuracy$California, 
                     x = 0.01, y = .6, 
                     width = inset.w, height = inset.h, vjust = 1) +
  cowplot::draw_plot(gg.accuracy$`Intermountain West`, 
                     x = .275, y = .4, 
                     width = inset.w, height = inset.h, vjust = 1) +
  cowplot::draw_plot(gg.accuracy$`Central America`,  
                     x = .775, y = .8, 
                     width = inset.w, height = inset.h, vjust = 1)

gg.asia.region.inset <- cowplot::ggdraw(gg.asia.region) +
  #Arrows
  cowplot::draw_grob(grid::grid.lines(x = c(.3, .4), y = c(.3, .6), 
                                      arrow=inset.arrow)) +
  cowplot::draw_grob(grid::grid.lines(x = c(.825, .8), y = c(.55, .4), 
                                      arrow=inset.arrow)) +
  # Inset plots
  cowplot::draw_plot(gg.accuracy$Himalayas,
                     x = .15, y = 0.4, 
                     width = inset.w, height= inset.h, vjust = 1) +
  cowplot::draw_plot(gg.accuracy$`Southeast Asia`,
                     x = .775, y = .935, 
                     width = inset.w, height = inset.h, vjust = 1)

# Fire
gg.america.burned.inset <- cowplot::ggdraw(gg.america.burned) +
  cowplot::draw_plot(gg.burned.fraction$`Pacific Northwest`, 
                     x = 0, y = 1, 
                     width = inset.w, height = inset.h, vjust = 1) +
  cowplot::draw_plot(gg.burned.fraction$California, 
                     x = 0, y = .6, 
                     width = inset.w, height = inset.h, vjust = 1) +
  cowplot::draw_plot(gg.burned.fraction$`Intermountain West`, 
                     x = .275, y = .4, 
                     width = inset.w, height = inset.h, vjust = 1) +
  cowplot::draw_plot(gg.burned.fraction$`Central America`, 
                     x = .775, y = .8, 
                     width = inset.w, height = inset.h, vjust = 1)

gg.asia.burned.inset <- cowplot::ggdraw(gg.asia.burned) +
  cowplot::draw_plot(gg.burned.fraction$Himalayas,
                     x = .15, y = 0.4,
                     width = inset.w, height= inset.h, vjust = 1) +
  cowplot::draw_plot(gg.burned.fraction$`Southeast Asia`,
                     x = .775, y = .935, 
                     width = inset.w, height = inset.h, vjust = 1)

# Precipitation
inset.w.p = .23
inset.h.p = .41
gg.america.precip.inset <- cowplot::ggdraw(gg.america.precip) +
  cowplot::draw_plot(gg.climatology$`Pacific Northwest`, 
                     x = 0, y = 1.01, 
                     width = inset.w.p, height = inset.h.p, vjust = 1) +
  cowplot::draw_plot(gg.climatology$California, 
                     x = .085, y = .64, 
                     width = inset.w.p, height = inset.h.p, vjust = 1) +
  cowplot::draw_plot(gg.climatology$`Intermountain West`, 
                     x = .245, y = .395, 
                     width = inset.w.p, height = inset.h.p, vjust = 1) +
  cowplot::draw_plot(gg.climatology$`Central America`, 
                     x = .71, y = .8, 
                     width = inset.w.p, height = inset.h.p, vjust = 1)

gg.asia.precip.inset <- cowplot::ggdraw(gg.asia.precip) +
  cowplot::draw_plot(gg.climatology$Himalayas,
                     x = .15, y = 0.4,  
                     width = inset.w.p, height= inset.h.p, vjust = 1) +
  cowplot::draw_plot(gg.climatology$`Southeast Asia`,
                      x = .7, y = .935, 
                     width = inset.w.p, height = inset.h.p, vjust = 1)
```

```{r map_panel, fig.height=8, fig.width=7}
panel.labels <- tibble(
  label = as_factor(c('Location accuracy distribution\nfor burned and unburned sites by region',
                      'Locations of burned sites and\nregional distribution of burned fraction',
                      str_c('Landslide-triggering precipitation percentile\nand',
                            'regional seasonal precipitation (mm)')))
)
gg.panel.labels <- ggplot(panel.labels) +
  facet_wrap(~label, ncol = 1, switch = T) +
  theme(panel.background = element_blank(),
        strip.text = element_text(face="bold", size=7))

gg.panel.maps <- cowplot::plot_grid(
  cowplot::plot_grid(gg.america.region.inset, gg.asia.region.inset,
                   labels = c('a', 'b'), label_x = 1, hjust = 1, vjust = 1),
  legend.regions,
  cowplot::plot_grid(gg.america.burned.inset, gg.asia.burned.inset,
                     labels = c('c', 'd'), label_x = 1, hjust = 1, vjust = 1),
  legend.burned,
  cowplot::plot_grid(gg.america.precip.inset, gg.asia.precip.inset,
                     labels = c('e', 'f'), label_x = 1, hjust = 1, vjust = 1),
  legend.precip,
  ncol = 1, rel_heights = c(1, .3, 1, .2, 1, .3))
gg.panel <- cowplot::plot_grid(gg.panel.labels, gg.panel.maps, 
                               nrow = 1, rel_widths = c(.5, 6.5))
gg.panel

ggsave(filename = "../figures/panel_map.png", width=7, height=8)
```


## Fire timing

This figure shows the timing of fires relative to landslides for all events that
took place at a burned site.
```{r fire_timing}
calc.fire.timing <- function (modis.date, modis, slide) {
  modis.date %>%
    dplyr::filter(burn_days_before >= 0) %>%
    dplyr::group_by(OBJECTID) %>%
    dplyr::filter(burn_days_before == min(burn_days_before))  %>%
    right_join(modis %>% dplyr::filter(burned=='burned'), by='OBJECTID') %>%
    inner_join(slide.regions, by='OBJECTID') %>%
    left_join(slide %>% dplyr::select(OBJECTID, event_yday, season), by='OBJECTID') %>% 
    dplyr::mutate(burn_yday = yday(burn_date),
                  burn_season = factor(burn_season,
                                       levels = c('Spring', 'Summer', 'Fall', 'Winter'),
                                       ordered = T))
}

get.data('fire.timing', overwrite = F,
         calc.fire.timing, 
         modis.date, modis, slide)
```

Most landslides take place within one year of the fire. California stands out as having over a 
quarter of the landslides taking place more than a year after the fire.
```{r calc_fire_timing_proportions}
fire.timing %>%
  dplyr::mutate(gt.1year = ifelse(burn_days_before > 2*365, 
                                  'gt.1year', 'lt.1year')) %>%
  dplyr::group_by(region, gt.1year) %>%
  dplyr::tally() %>%
  tidyr::spread(gt.1year, n, drop=F) %>%
  dplyr::mutate(prop.gt.1year = gt.1year / (gt.1year + lt.1year))
  
```

```{r plot_fire_landslide_timing, fig.height = 8, fig.width = 6.5}
fire.to.landslide.x.breaks <- seq(season.breaks[4] - 4 * 365, season.breaks[3], 
                                  365/4)
fire.to.landslide.x.labels <- map(c(str_c('-', c(3, 2, 1), 'y '), ''), 
                                  ~c(str_c(., 'Winter'), 
                                     'Spring', 'Summer', 'Fall')) %>% as_vector

fire.facet.labels <- fire.timing %>%
  dplyr::group_by(region) %>%
  dplyr::mutate(wrap_event_yday = ifelse(event_yday > season.breaks['winter'], 
                                         event_yday - 365, event_yday)) %>%
  dplyr::ungroup() %>%
  dplyr::mutate(ymin = min(wrap_event_yday - burn_days_before),
                ymax = max(wrap_event_yday)) %>%
  dplyr::group_by(region) %>%
  dplyr::summarize(y = n() * 0.8, x = min(ymin) + (max(ymax) - min(ymin)) * 0.1) %>%
  dplyr::mutate(label = c('a', 'b', 'c', 'd', 'e', 'f', 'g'))

fire.timing %>%
  dplyr::group_by(region) %>%
  dplyr::mutate(wrap_event_yday = ifelse(event_yday > season.breaks['winter'], 
                                         event_yday - 365, event_yday)) %>%
  dplyr::arrange(wrap_event_yday - burn_days_before) %>%
  dplyr::mutate(i = row_number()) %>%
  ggplot() +
  facet_wrap(vars(region), scales = 'free_y', ncol = 2, shrink=F) +
  geom_segment(aes(x = wrap_event_yday - burn_days_before, xend = wrap_event_yday,
                   y = i, yend = i, 
                   color = burn_season)) +
  geom_point(aes(x = wrap_event_yday, y = i, shape='Landslide'), size=0.5) +
  geom_vline(data = tibble(x=seq(334-365*3, 365, 365)), 
             aes(xintercept = x, size = 'start_of_winter')) +
  geom_vline(aes(xintercept = 334 - 365, size = 'year_of')) +
  geom_rug(aes(y = i, color = burn_season)) +
  geom_rug(aes(x = event_yday - burn_days_before), sides = 't', length = unit(0.05, "npc")) +
  geom_label(data = fire.facet.labels,
             aes(x = x, y = y, label = label)) +
  theme_bw() +
  scale_x_continuous('Landslides and preceding fires relative to the year of the landslide',
                     minor_breaks = NULL,
                     breaks = as_vector(fire.to.landslide.x.breaks),
                     labels = fire.to.landslide.x.labels) +
  scale_color_manual('Fire season:',
                     values = c('Spring'='#19E667', 'Summer'='#E61998', 
                                'Fall'='#E6CD19', 'Winter'='#1932E6'),
                     guide = guide_legend(nrow=1, order = 2)) + 
  scale_y_continuous('Post-wildfire landslides ordered by fire timing relative to landslide', 
                     expand = c(0.1, 0.1)) +
  scale_size_manual('Relative timing:',
                    values = c('year_of' = 1, 'start_of_winter' = 0.5),
                    labels = c('year_of' = 'Start of year of landslide',
                               'start_of_winter' = 'Start of winter (December 1)'),
                    guide = guide_legend(nrow=2, order = 3)) +
  guides(shape = guide_legend(title = NULL, override.aes = c(size = 1.5), order = 1)) +
  theme(panel.grid.major.x = element_line(color = 'grey', size = 0.5),
        axis.text.x = element_text(hjust = 1, vjust = 1, angle = 90),
        axis.ticks = element_blank(),
        axis.text.y = element_blank(),
        legend.position = c(.75, .04),
        legend.direction = 'vertical',
        legend.box = 'vertical',
        legend.margin = unit(5, 'pt'),
        legend.box.margin = margin(5, 5, 5, 5),
        legend.box.background = element_rect(color = 'black', size = 2),
        legend.key.height = unit(10, 'pt') 
  )
ggsave('../figures/fire_to_landslide.png', width = 6.5, height = 8)
```


# Analysis 

## CHIRPS vs Daymet

To validate the exclusion of zero-precipitation landslides, a comparison of CHIRPS and Daymet
in California and Nevada only. Many of the zero-precipitation landslides measured by CHIRPS are
not the same as the ones measured by Daymet, suggesting that these are measurement errors. 

Approximately 14% of the landslides in California and Nevada were excluded on this basis.

```{r join_chirps_daymet}
join.chirps.daymet <- function (precip.chirps, precip.daymet) {
  precip.chirps %>%
    dplyr::inner_join(precip.daymet %>% dplyr::select(-days_before, -yday), 
                      by=c('OBJECTID', 'date', 'window'), 
                      suffix=c('.chirps', '.daymet'))
}

get.data('precip.chirps.daymet.ca.event', join.chirps.daymet, overwrite = F,
         precip.chirps.ca.event, 
         precip.daymet.ca.all)
get.data('precip.chirps.daymet.ca.all', join.chirps.daymet, overwrite = F,
         precip.chirps.ca.all, precip.daymet.ca.all)
precip.chirps.ca.event %>%
  filter(days_before==-1, window==7, !is.na(pctl)) %>%
  group_by(precip = ifelse(pctl > 0, 'non.zero.precip', 'zero.precip')) %>%
  tally() %>%
  spread(key = precip, value = n) %>%
  mutate(total = non.zero.precip + zero.precip,
         percent.zero.precip = zero.precip / total)
```


```{r plot_chirps_daymet, fig.height=4, fig.width=6}
precip.chirps.daymet.ca.all %>%
  tidyr::drop_na(pctl.chirps, pctl.daymet) %>%
  ggplot() +
  geom_point(aes(x=pctl.chirps, y=pctl.daymet, color = 'all'), 
             alpha=0.2,shape = 16) +
  geom_point(data = precip.chirps.daymet.ca.event %>% 
               dplyr::filter(window == 7, days_before == -1 ) %>%
               dplyr::mutate(to.remove = ifelse(pctl.chirps==0, 'exclude', 'keep')),
             aes(x=pctl.chirps, y=pctl.daymet, color=to.remove)) +
  coord_equal() +
  theme_bw() +
  scale_color_manual('', 
                     values = c('keep'='black', 'exclude'='blue', 'all'='gray'),
                     labels = c('keep'='Included\nday-of-landslide\npreciptiation',
                                'exclude'='Excluded\nday-of-landlside\nprecipitation',
                                'all'='Precipitation\nfrom all dates')) +
  guides(color = guide_legend(keyheight = grid::unit(4, 'char'))) +
  labs(x='CHIRPS 7-day Precipitation Percentile',
       y='Daymet 7-day Precipitation Percentile')

ggsave('../figures/chirps_vs_daymet.png', width=6, height=4, dpi=600)
```


## Precipitation percentile wilcox test

Comparison of antecedent precipitation time series among burned and unburned sites by region.
Mann-Whitney tests are used to establish whether or not differences are statistically significant.

```{r wilcox_functions}
calc.burn.wilcox <- function (precip, pre=30, post=10, alternative='two.sided', 
                              source=NA, groupby=c('burned', 'days_before')) {
  precip %>%
    dplyr::filter(days_before <= pre, days_before >= -post) %>%
    dplyr::group_by_at(groupby) %>%
    calc.wilcox(alternative, source)
}
  
calc.wilcox <- function (precip, alternative='two.sided', source=NA) {
  wilcox <- precip %>%
    tidyr::nest() %>%
    spread(key = 'burned', value = 'data') %>%
    dplyr::mutate(wilcox = map2(burned, unburned, 
                         ~wilcox.test(.x$pctl, .y$pctl, 
                                      alternative=alternative))) %>%
    dplyr::mutate(wilcox.glance = map(wilcox, glance)) %>%
    tidyr::unnest(wilcox.glance) %>%
    dplyr::mutate(signif.95 = p.value <= 0.05) %>%
    gather(key = 'burned', value = 'data', burned, unburned) %>%
    tidyr::unnest(data)
  if (!is.na(source)) wilcox <- wilcox %>% dplyr::mutate(label=source)
  wilcox
}

magnitude.regional.data <- function (slide.regions, precip, 
                                     pre = 30, post = 10, w = 7) {
  precip %>%
    dplyr::filter(days_before < pre, days_before > -post, window==w) %>%
    right_join(slide.regions, by = 'OBJECTID') %>%
    dplyr::left_join(modis, by = 'OBJECTID') %>%
    dplyr::filter(pctl > 0) %>%
    droplevels() %>%
    dplyr::select(region, burned, days_before, pctl) %>%
    calc.burn.wilcox(pre = pre, post = post, alternative = 'less',
                     groupby = c('region', 'burned', 'days_before'))
}

get.data('burn.wilcox.regions.event', overwrite = F,
         magnitude.regional.data,
         slide.regions,
         precip.chirps.ww.event)
```

```{r plot_burn_wilcox, fig.height=11, fig.width=11}
facet.levels <- c('Rolling cumulative\nprecipitation percentile',
                  'Mann-Whitney p-value')

plot.burn.wilcox <- function (burn.wilcox.regions.event, slide.regions) {
  region.tally <- slide.regions %>% 
    left_join(modis) %>% 
    dplyr::group_by(burned=burn.cumulative > 0, region) %>% 
    tally
  
  burn.wilcox.regions.event %>%
    dplyr::mutate(signif.95 = ifelse(signif.95, 'significant', 'not.significant')) %>%
    ggplot() +
    facet_wrap(region~., ncol=1, strip.position = 'right') +
    geom_boxplot(aes(x = days_before, 
                     group=interaction(days_before, burned), 
                     y = pctl, 
                     fill = burned,
                     alpha = interaction(signif.95, burned)), 
                 outlier.shape=21)  +
    annotate("rect", xmin = -0.5, xmax = 0.5, ymin = 0, ymax = 1, alpha = .4) +
    geom_label(data=region.tally, 
               aes(x=-15, y=0.1+0.25*burned, 
                   label=str_c(ifelse(burned, 'Burned: ', 'Unburned: '), n)),
               hjust='right', label.size = 0) +
    geom_label(data = tibble(region = region.plot.order,
                             label = c('a', 'b', 'c', 'd', 'e', 'f', 'g')) %>%
                 dplyr::mutate(region = factor(region, region.plot.order, ordered = T)),
               aes(x = -15, y = 0.85, label = label), hjust = 'right') +
    scale_x_reverse(limits = c(31, -16), expand = expansion()) +
    scale_fill_manual('', 
                      values = c('burned' = '#d95f02', unburned = '#7570b3'),
                      labels = c('burned'='Burned', 'unburned'='Unburned'),
                      guide = guide_legend(direction = 'horizontal')) +
    scale_alpha_manual('Statistical significance at\n95% confidence',
                       values=c('not.significant.burned' = 0.4, 
                                'not.significant.unburned' = 0.4,
                                'significant.burned' = 1, 
                                'significant.unburned' = 1),
                       labels=c('not.significant.burned' = '', 
                                'not.significant.unburned' = 'Not Significant',
                                'significant.burned' = '', 
                                'significant.unburned' = 'Significant'),
                       breaks=c('not.significant.burned', 
                                'not.significant.unburned',
                                'significant.burned', 
                                'significant.unburned'),
                       guide=guide_legend(override.aes=list(fill=c('not.significant.burned' = '#d95f02', 
                                                                   'not.significant.unburned' = '#7570b3',
                                                                   'significant.burned' = '#d95f02', 
                                                                   'significant.unburned' = '#7570b3')),
                                          direction = 'horizontal')) +
    theme_bw() +
    labs(x='Days before the event', y='Rolling cumulative precipitation percentile') +
    theme(legend.position = 'bottom', legend.direction = 'vertical')
}

burn.wilcox.regions.event %>%
  plot.burn.wilcox(slide.regions)

ggsave('../figures/magnitude_regional.png', width = 11, height = 11)
```


## Precipitation (mm) bootstrap sampling

Bootstrap samples are used here to develop baseline precipitation values against which to compare
landslide-triggering precipitation. This method is an improvement over the original Wilcox tests because
sample-size disparities between regions are reduced. 

### Bootstrap sampling

```{r bootstrap_functions}
sample.bootstrap <- function(precip.subset, precip.all, n, samples, burned, region) {
  precip.all %>%
    dplyr::filter(burned==burned, region==region) %>%
    dplyr::mutate(precip.yday = yday(date)) %>%
    # dplyr::select precipitation from same location and date of year as subset
    dplyr::inner_join(precip.subset %>% 
                        dplyr::select(OBJECTID, event_yday), 
                      by=c('OBJECTID'='OBJECTID', 
                           'precip.yday'='event_yday')) %>%
    # Sample precipitation matching the site/day in bulk
    dplyr::group_by(OBJECTID) %>%
    tidyr::drop_na(pctl) %>%
    dplyr::sample_n(n, replace = T) %>%
    # Randomly split into individual samples
    dplyr::mutate(i = sample(1:samples, n, replace = T )) 
}

calc.burn.bootstrap <- function (precip.event, precip.all, 
                                 samples=100, per.subset=10000,
                                 pre=30, post=10, w=7) {
  groupby = c('region', 'burned', 'days_before')
  precip.all <- precip.all %>% 
    dplyr::filter(window==w) %>% 
    dplyr::select(-window)
  precip.event %>%
    rename(precip.date = date, precip.yday = yday) %>%
    dplyr::filter(days_before <= pre, days_before >= -post, window==w) %>%
    dplyr::select_at(c(groupby, 'OBJECTID', 'event_yday')) %>%
    # Tally each subset/site/day combination
    dplyr::group_by_at(groupby) %>%
    add_tally(name = 'sites.in.subset') %>%
    # Determine number to draws from each site that will 
    # roughly match sample sizes among subsets
    # Round up - number of draws must be an integer
    dplyr::mutate(n.per.site = ceiling(per.subset / sites.in.subset)) %>%
    # Take bootstrap samples for each subset/site/day combination
    dplyr::group_by_at(c(groupby, 'n.per.site', 'sites.in.subset')) %>%
    tidyr::nest() %>%
    dplyr::mutate(bootstrap.samples = pmap(list(data, n.per.site, 
                                                burned, region), 
                                           ~sample.bootstrap(..1, precip.all, 
                                                             ..2, samples, 
                                                             ..3, ..4)))
}
```

```{r bootstrap_sample_regions}
join.details <- function (precip, slide, slide.regions, modis) {
  precip %>% 
    dplyr::inner_join(slide %>% dplyr::select(OBJECTID, event_yday), 
                      by='OBJECTID') %>%
    dplyr::right_join(slide.regions, by = 'OBJECTID') %>%
    dplyr::left_join(modis %>% dplyr::select(OBJECTID, burned), 
                     by='OBJECTID')
}

get.bootstrap <- function (reg='All landslides', 
                           name = 'precip.chirps.bootstrap.all', 
                           ovw=F) {
  get.data(name, overwrite = ovw,
           calc.burn.bootstrap,
           precip.chirps.ww.event %>% 
             join.details(slide, slide.regions %>% filter(region == reg), modis),
           precip.chirps.ww.all %>% 
             join.details(slide, slide.regions %>% filter(region == reg), modis))
}

get.bootstrap('California', 'precip.chirps.bootstrap.cal', ovw = F)
get.bootstrap('Intermountain West', 'precip.chirps.bootstrap.imw', ovw = F)
get.bootstrap('Pacific Northwest', 'precip.chirps.bootstrap.pnw', ovw = F)
get.bootstrap('Central America', 'precip.chirps.bootstrap.cam', ovw = F)
get.bootstrap('Himalayas', 'precip.chirps.bootstrap.him', ovw = F)
get.bootstrap('Southeast Asia', 'precip.chirps.bootstrap.sea', ovw = F)
get.bootstrap(ovw = F)

precip.chirps.bootstrap <- precip.chirps.bootstrap.all %>%
  dplyr::bind_rows(precip.chirps.bootstrap.cal) %>%
  dplyr::bind_rows(precip.chirps.bootstrap.imw) %>%
  dplyr::bind_rows(precip.chirps.bootstrap.pnw) %>%
  dplyr::bind_rows(precip.chirps.bootstrap.cam) %>%
  dplyr::bind_rows(precip.chirps.bootstrap.him) %>%
  dplyr::bind_rows(precip.chirps.bootstrap.sea)

precip.chirps.bootstrap %>%
  dplyr::filter(days_before==0) %>%
  dplyr::select(-data) %>%
  dplyr::mutate(site.total = map_int(bootstrap.samples, ~nrow(ungroup(.))),
                n.per.site.real = site.total / sites.in.subset,
                avg.sample.size = map_dbl(bootstrap.samples, 
                                          ~group_by(., i) %>% 
                                            summarize(n = n()) %>% 
                                            pull(n) %>% mean)) %>%
  dplyr::select(-bootstrap.samples)
```

### Summarize bootstrap samples
```{r bootstrap_mann_whitney, class.source = 'fold.show'}
mann.whitney.p.value <- function (subset, bootstrap) {
  bootstrap %>%
    dplyr::group_by(i) %>%
    tidyr::nest() %>%
    # test if bootstrap samples are less than pre-landslide precipitation
    dplyr::mutate(wilcox = map(data,
                               ~wilcox.test(pull(., pctl), 
                                            pull(subset, pctl), 
                                            alternative='less'))) %>% 
    dplyr::mutate(wilcox.glance = map(wilcox, glance)) %>%
    dplyr::select(-data, -wilcox) %>%
    tidyr::unnest(wilcox.glance)
}
```

```{r subset_precip}
subset.precip <- function (precip, regions, modis, w=7, pre=30, post=10) {
  precip %>%
    dplyr::filter(window==7, days_before <= pre, days_before >= -post) %>%
    dplyr::select(OBJECTID, pctl, days_before) %>%
    dplyr::right_join(regions, by='OBJECTID') %>%
    dplyr::inner_join(modis, by='OBJECTID') %>%
    dplyr::group_by(region, burned, days_before) %>%
    tidyr::nest()
}

get_subset <- function (reg, bur, day, subsets) {
  row <- subsets %>%
    dplyr::filter(region==as.character(reg),
                  burned==bur,
                  days_before==day)
  row$data[[1]]
}

get.data('precip.chirps.ww.event.subsets', 
         subset.precip,
         precip.chirps.ww.event,
         slide.regions,
         modis)
```

```{r summarize_bootstrap} 
summarize.bootstrap <- function (bootstrap.df, subsets, objective.function) {
  bootstrap.df %>%
    dplyr::mutate(objective = purrr::pmap(list(region, 
                                               burned, 
                                               days_before,
                                               bootstrap.samples), 
                                          ~objective.function(get_subset(..1, 
                                                                         ..2, 
                                                                         ..3, 
                                                                         subsets),
                                                              ..4))) %>%
    dplyr::select(-bootstrap.samples) %>%
    tidyr::unnest(objective) %>%
    dplyr::ungroup()
}
```



```{r summarize_bootstrap_wilcox}
get.data('precip.chirps.bootstrap.wilcox', overwrite = F,
         summarize.bootstrap,
         precip.chirps.bootstrap %>% dplyr::select(-data),
         precip.chirps.ww.event.subsets,
         mann.whitney.p.value)
precip.chirps.bootstrap.wilcox %>% 
  dplyr::select(region, burned, days_before, p.value)
```


The bootstrap plot also shows the day-of-landslide kernel density estimates for all bootstrap
samples against the landslide-triggering precipitation to illustrate the method.
### Plot bootstrap sampling
```{r bootstrap_wilcox, fig.width=11, fig.height=11, fig.show='hide'}
plot.burn.bootstrap <- function (bootstrap, var, name) {
  bootstrap %>%
    ggplot()  +
    facet_wrap(.~region, ncol=1, strip.position='right') +
    geom_vline(aes(xintercept=0)) +
    geom_boxplot(aes(x = days_before, 
                     group = interaction(days_before, burned), 
                     y = {{ var }}, 
                     fill = burned,
                     color = burned),
                 outlier.shape=21,
                 alpha = 0.6,
                 position = position_dodge(width=0.4, preserve = 'total'))  +
    annotate("rect", xmin = -11, xmax = -0.1, ymin = 0.05, ymax = .9999, 
             fill = 'white', alpha = 0.4) +
    annotate("rect", xmin = 0.1, xmax = 31, ymin = 0.05, ymax = .9999, 
             fill = 'white', alpha = 0.4)  +
    geom_hline(aes(yintercept=0.05), linetype='dashed') +
    geom_label(data = tibble(region = region.plot.order,
                             label = c('a', 'b', 'c', 'd', 'e', 'f', 'g')) %>%
                 dplyr::mutate(region = factor(region, region.plot.order, ordered = T)),
               aes(x = -11.5, y = 0.9999, label = label), 
               hjust = 'right', vjust = 'bottom')  +
    scale_x_reverse(expand = expansion(), limits = c(31, -12)) +
    scale_y_continuous(trans = probit_trans(),
                       breaks = bootstrap.y.breaks,
                       labels = bootstrap.y.labels) +
    scale_fill_burned +
    scale_color_burned +
    coord_trans(y = 'reverse', 
                ylim = c(10^-35, .9999)) +
    theme_bw() +
    labs(x='Days before the landslide', y=name) +
    theme(legend.position = 'bottom',
          panel.grid.minor.y = element_blank())
}

bootstrap.y.breaks <- c(10^-30,  10^-10,  0.05,   0.5,   0.9999)
bootstrap.y.labels <- c('1e-30', '1e-10', '0.05', '0.5', '1')
plot.bootstrap.wilcox <- precip.chirps.bootstrap.wilcox %>%
  plot.burn.bootstrap(p.value, 
                      'p-value of Wilcox tests between bootstrap samples and pre-landslide precipitation') 

plot.bootstrap.wilcox
```

```{r sample_kde, fig.width = 4, fig.height = 11, fig.show='hide'}
plot.bootstrap.sample.kde <- precip.chirps.bootstrap %>%
  dplyr::filter(days_before==0) %>%
  dplyr::select(-data) %>%
  tidyr::unnest() %>%
  ggplot() +
  facet_grid(cols = vars(burned), rows = vars(region), scales = 'free_y') +
  geom_density(aes(x = pctl, group = i, color = burned, fill = burned), 
               alpha = 0.01) +
  geom_density(data = precip.chirps.ww.event %>% 
                 filter(days_before==0) %>% 
                 right_join(slide.regions, by = 'OBJECTID')%>%
                 inner_join(modis, by='OBJECTID'),
               aes(x = pctl, color = 'day_of', fill = 'day_of'), size = 1) +
  geom_label(data = cross_df(list(region = region.plot.order, 
                                  burned = c('burned', 'unburned'))) %>%
               dplyr::mutate(label = c('h', 'i', 'j', 'k', 'l', 'm', 'n', 
                                       'o', 'p', 'q', 'r', 's', 't', 'u')) %>%
               dplyr::mutate(region = factor(region, region.plot.order, ordered = T)),
             aes(x = .95, y = .05, label = label), 
             hjust = 'right', vjust = 'bottom') +
  scale_fill_manual('',
                    breaks = c('day_of', 'burned', 'unburned'),
                    values = c('day_of' = NA, burned.colors),
                    labels = c('day_of' = 'Day of landslide precipitation',
                               'burned' = 'Bootstrap samples, burned sites',
                               'unburned' = 'Bootstrap samples, unburned sites'),
                    guide = guide_legend(override.aes = list(alpha = 0.5))) +
  scale_color_manual('', 
                     breaks = c('day_of', 'burned', 'unburned'),
                     values = c('day_of' = 'black', burned.colors),
                     labels = c('day_of' = 'Day of landslide precipitation',
                                'burned' = 'Bootstrap samples, burned sites',
                                'unburned' = 'Bootstrap samples, unburned sites'),
                    guide = guide_legend(override.aes = list(alpha = 0.5))) +
  scale_x_continuous(breaks = seq(0, 1, 0.25),
                     labels = c('0', '0.25', '0.5', '0.75', '1')) +
  labs(y = 'Kernel density estimate', x = 'Precipitation percentile') +
  theme_bw() +
  theme(legend.position = 'bottom',
        legend.spacing.x = unit(15, "pt"),
        axis.text.y = element_blank(),
        axis.text.x = element_text(),
        axis.ticks.y = element_blank(),
        strip.text.x = element_blank(),
        panel.spacing.x = unit(15, 'pt')) +
  coord_cartesian(expand = F)
plot.bootstrap.sample.kde
```

```{r plot_bootstrap_wilcox_kde, fig.width = 13, fig.height = 11}
plot.bootstrap.sample.kde.legend <- get_legend(plot.bootstrap.sample.kde)
plot.bootstrap.wilcox.legend <- get_legend(plot.bootstrap.wilcox)
cowplot::plot_grid(cowplot::plot_grid(plot.bootstrap.wilcox + 
                      theme(strip.text = element_blank(),
                            plot.margin = unit(c(5,5,5,5), 'pt'),
                            legend.position = 'none',
                            legend.direction = 'horizontal'), 
                    plot.bootstrap.sample.kde + 
                      theme(plot.margin = unit(c(5,5,5,2), 'pt'),
                            legend.position = 'none'),
                    rel_widths = c(11, 4)),
          cowplot::plot_grid(plot.bootstrap.wilcox.legend, 
                    plot.bootstrap.sample.kde.legend,
                    rel_widths = c(3, 7)),
          nrow = 2, rel_heights = c(10, 0.5))

ggsave('../figures/precip_bootstrap_wilcox+kde.png', width=13, height=11)
```


## Seasonal density and precipitation frequency

### Seasonal density

The kernel density estimate of the annual distribution of landslides is used to show
shifts in seasonality relative to the day of the year.

```{r seasons_by_region_polar, fig.width=2.5, fig.height=11, fig.show='hide'}
slide.season <- slide %>%
  dplyr::select(OBJECTID, location_accuracy, season, event_yday) %>%
  dplyr::right_join(slide.regions, by='OBJECTID') %>%
  dplyr::left_join(modis, by='OBJECTID')

plot.seasonality <- bind_rows(slide.season %>% 
                                dplyr::mutate(event_yday = event_yday - 365),
                              slide.season,
                              slide.season %>% 
                                dplyr::mutate(event_yday = event_yday + 365)) %>%
  dplyr::group_by(region, burned) %>%
  tidyr::drop_na(burned) %>%
  tidyr::nest() %>%
  dplyr::mutate(kde = map(data, 
                          function (df) density(df$event_yday, 
                                                from=-100, to=450, 
                                                n = 512, bw = 21)),
                kde = map(kde, function (dens) tibble(event_yday = dens$x, 
                                                      density = dens$y))) %>%
  tidyr::unnest(kde) %>%
  dplyr::filter(event_yday >= 0, event_yday < 365) %>%
  ggplot() +
  facet_wrap(region~., ncol = 1, strip.position = 'right') +
  geom_area(aes(x = event_yday, y = density, group = burned, fill=burned),
            trim = T, alpha = 0.6, color = 'black', 
            position = position_identity()) +
  geom_text(data = tibble(text = c('h', 'i', 'j', 'k', 'l', 'm', 'n'),
                          hjust = c(-3.5, -10, -10, -4, -10, -2, -3.5),
                          region = region.plot.order,
                          x = -2.1*365, y = 0.50) %>%
               mutate(region = factor(region, region.plot.order, ordered = T)),
            aes(x = 0, y = .0035, label = text, hjust = hjust), size = 7) +
  coord_polar() +
  scale_fill_burned +
  scale_x_continuous(breaks = (season.breaks + 365/8) %% 365,
                     labels = c('W', 'Sp', 'Su', 'F'),
                     minor_breaks = season.breaks) +
  labs(y = '', x = ' ', fill = 'Season') +
  theme_bw() +
  theme(axis.text.y = element_blank(),
        axis.ticks.y = element_blank(),
        axis.text.x = element_text(angle = 0, vjust = 1),
        legend.position = 'bottom',
        panel.grid.minor.x = element_line(color = 'black'))

plot.seasonality
```

### Precipitation frequency

Smoothed precipitation frequency is used to show shifts in the annual distribution
of landslides relative to the annual precipitation seasonality. 

```{r calc_regional_frequency}
n.years <- 2
calc.prop <- function (col) sum(!(col %in% 0)) / n()
calc.yday.frequency <- function (precip, slide, slide.regions, modis) {
  precip %>%
  dplyr::mutate(precip_yday = yday(date)) %>% 
  dplyr::select(OBJECTID, precip_yday, pctl) %>%
  dplyr::inner_join(slide %>% dplyr::select(OBJECTID, event_yday), 
                    by='OBJECTID') %>%
  dplyr::mutate(ydays_before = event_yday - precip_yday) %>%
  dplyr::select(OBJECTID, ydays_before, pctl) %>%
  dplyr::mutate(ydays_before = ydays_before %% 365) %>%
  dplyr::right_join(slide.regions, by = 'OBJECTID') %>%
  dplyr::left_join(modis %>% dplyr::select(OBJECTID, burned), 
                   by='OBJECTID') %>%
  dplyr::group_by(region, burned, ydays_before) %>%
  dplyr::summarize_at(vars(pctl), list(precip.yday.freq = calc.prop))
}
 
get.data('precip.chirps.yday.frequency',
         calc.yday.frequency,
         precip.chirps.ww.all,
         slide,
         slide.regions,
         modis)

calc.frequency <- function (precip, precip.yday.freq, 
                            slide, slide.regions, modis) {
  precip.chirps.ww.all %>%
    dplyr::transmute(OBJECTID, date = lubridate::as_date(date), pctl) %>%
    dplyr::inner_join(slide %>% dplyr::select(OBJECTID, event_date), 
                      by='OBJECTID') %>%
    dplyr::mutate(days_before = as.integer(event_date - date)) %>%
  dplyr::select(OBJECTID, days_before, pctl) %>%
    dplyr::filter(days_before < n.years * 365, 
                  days_before > -n.years * 365) %>%x
    dplyr::right_join(slide.regions, by = 'OBJECTID') %>%
    dplyr::left_join(modis %>% dplyr::select(OBJECTID, burned), 
                     by='OBJECTID') %>%
    dplyr::group_by(region, burned, days_before) %>%
    dplyr::summarize_at(vars(pctl), list(precip.day.freq = calc.prop)) %>%
    dplyr::mutate(ydays_before = days_before %% 365) %>%
    dplyr::left_join(precip.yday.freq, by=c('region', 'ydays_before', 'burned')) %>%
    dplyr::mutate(precip.freq.norm  = precip.day.freq / precip.yday.freq,
                  precip.freq.anomaly = precip.day.freq - precip.yday.freq,
                  precip.freq.lt.anomaly = precip.day.freq - mean(precip.day.freq)) %>%
    dplyr::ungroup()
}
get.data('precip.chirps.frequency', overwrite = F,
         calc.frequency,
         precip.chirps.ww.all,
         precip.chirps.yday.frequency,
         slide,
         slide.regions,
         modis)
```

```{r precip_frequency, fig.width=10, fig.height=12, fig.show='hide'}
plot.frequency <- function (data, column, y.lable, n.years, groupby=c('burned')) {
  data %>%
    dplyr::group_by_at(groupby) %>%
    dplyr::mutate(smooth = rollmean(!!as.name(column), 90, na.pad=TRUE)) %>%
    ggplot() +
    geom_line(aes(x = days_before, y = !!as.name(column), 
                  group = burned, color = burned, size = 'daily')) +
    geom_line(aes(x = days_before, y = smooth, group = burned, color = burned, 
                  size = 'rolling')) +
    geom_hline(data = data %>% 
                 dplyr::group_by_at(groupby) %>% 
                 summarize(mean = mean(!!as.name(column))),
               aes(yintercept = mean, color=burned), linetype = 'dashed') +
    geom_text(data = tibble(text = c('a', 'b', 'c', 'd', 'e', 'f', 'g'),
                            region = region.plot.order,
                            x = -1.95*365, y = 0.50),
              aes(x = x, y = y, label = text), size = 7,
              hjust = 'right') +
    scale_size_manual('',
                      values = c('daily' = 0.1, 'rolling' = 0.75),
                      labels = c('daily' = 'Daily frequency', 
                                 'rolling' = '90-day rolling average')) +
    scale_color_manual('', 
                      values = c('burned' = '#d95f02', unburned = '#7570b3'),
                      labels = c('burned'='Burned', 'unburned'='Unburned'),
                      guide = guide_legend(override.aes = list(size = 2))) +
    scale_x_reverse('Years before the landslide', 
                    breaks = seq(-365 * n.years, 365 * n.years, 365),
                    labels = function (breaks) breaks / 365,
                    expand = expansion()) +
    labs(y = y.lable) +
    theme_bw() +
    theme(legend.position = 'bottom')
}

plot.frequency.region <- precip.chirps.frequency %>%
  plot.frequency('precip.freq.lt.anomaly', 
                 'Precipitation frequency anomaly - long-term mean',
                 n.years=n.years, 
                 groupby=c('burned', 'region')) +
  facet_wrap(region~., ncol=1, strip.position = 'right')
plot.frequency.region
```

```{r plot_precip_frequency_seasonality, fig.width=12.5, fig.height=12}
cowplot::plot_grid(
  cowplot::plot_grid(
    plot.frequency.region + 
      theme(strip.text.y = element_blank(),
            legend.position = 'none',
            plot.margin = unit(c(5, 1, 5, 5), 'pt')),
    plot.seasonality +
      theme(legend.position = 'none',
            plot.margin = unit(c(5, 5, 12, 0), 'pt')),
    rel_widths = c(10, 2.5)
  ),
  get_legend(plot.frequency.region),
  nrow = 2, rel_heights = c(10, 0.5)
)

ggsave('../figures/precip_frequency_seasonality.png', width=11, height=12)
```


